Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)

### 2014-02-26 WSSSPE1_report/WSSSPE1_summary_report.tex
add note
minor changes

One of the tasks of catalysts is to seek community collaborations for new
applications and user engagement thus benefiting both: science, by effective
running of scientific codes on computational infrastructure, and technology,
by conducting research and seeking findings for technology improvement.

The particular engagements described in the paper came up from authors work as
postdoctoral researcher at Cornell and Argonne. Interaction with the scientific
communities in both institutions resulted in these collaborations.

@@

Nancy Wilkins-Diehr summarized 6 papers in ``Communities Part 2.'' The Christopherson paper outlines the degree to which research
relies on high quality software. There are often barriers and a lack of suitable incentives for researchers to embrace software engineering principals.
The Water Science Software Institute is working to lower some of these barriers through an Open Community Engagement Process.
This is a 4-step iterative development process that incorporates Agile development principals.

\item Step 1 - Design - thorough discussion of research questions
\item Step 2 - Develop working code
\item Step 3 - Refine based on new requirements
\item Step 4 - Publish open source

The Christopherson paper reports on the application of Steps 1-3 to a computational modeling framework developed in the 1990s.
Step 1 was realized as a 2-day, in person specifications meeting and code walkthrough. Step 2 was a 5-day hackathon to develop working code.
Step 3 was a 3-day hackathon to refine the code based on new requirements. The team worked on small, low-risk unit of
code. It was challenging, revealed unanticipated obstacles, programmers had to work together, and experimentation was encouraged.

The paper summarized recommendations to others wishing to engage in this or a similar process. Start small and gradually
build toward more complex objectives. This is consistent with Agile development. Refactor before adding new functionality.
Approach development as a learning experience. Welcome experimentation, and treat mistakes as a natural part of the learning process.
Repeat Step 1 activities before all hackathons to develop consensus before coding. This allows coding to be the focus of subsequent hackathons.
In higher risk situations, provide additional time for Step 1 activities. The Christopherson team recommends a minimum of two months.
Ensure any newcomers receive some form of orientation prior to the hackathon, such as a code walkthrough or system documentation.
Co-locate rather than collaborating remotely whenever feasible.

>>>>

One of the tasks of catalysts is to seek community collaborations for new
applications and user engagement thus benefiting both: science, by effective
running of scientific codes on computational infrastructure, and technology,
by conducting research and seeking findings for technology improvement.
The particular engagements described in the paper came up from authors work as
postdoctoral researcher at Cornell and Argonne. Interaction with the scientific
communities in both institutions resulted in these collaborations.

\note{More papers are needed here.}

@@

Nancy Wilkins-Diehr summarized 6 papers in this part.

The Christopherson paper outlines the degree to which research relies on high
quality software. There are often barriers and a lack of suitable incentives
for researchers to embrace software engineering principals. The Water Science
Software Institute is working to lower some of these barriers through an Open
Community Engagement Process. This is a 4-step iterative development process
that incorporates Agile development principals.

\begin{itemize}
\item Step 1: Design - thorough discussion of research questions
\item Step 2: Develop working code
\item Step 3: Refine based on new requirements
\item Step 4: Publish open source
\end{itemize}

The Christopherson paper reports on the application of Steps 1-3 to a
computational modeling framework developed in the 1990s. Step 1 was realized as
a 2-day, in person specifications meeting and code walkthrough. Step 2 was a
5-day hackathon to develop working code, and Step 3 a 3-day hackathon to refine
the code based on new requirements. The team worked on small, low-risk units of
code. It was challenging, revealed unanticipated obstacles, programmers had to
work together, and experimentation was encouraged.

The paper summarized recommendations to others wishing to engage in this or a
similar process, starting small and gradually building toward more complex
objectives. This is consistent with Agile development. Refactor before adding
new functionality. Approach development as a learning experience. Welcome
experimentation, and treat mistakes as a natural part of the learning process.
Repeat Step 1 activities before all hackathons to develop consensus before
coding. This allows coding to be the focus of subsequent hackathons. In higher
risk situations, provide additional time for Step 1 activities. The
Christopherson team recommends a minimum of two months. Ensure any newcomers
receive some form of orientation prior to the hackathon, such as a code
walkthrough or system documentation. Co-locate rather than collaborating
remotely whenever feasible.
